---
title: "Class 4 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

#library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview

 * Run a hypothesis test (randomization test) for a single proportion in R
 





## Part 1: Hypothesis tests for a single proportion 


Let us use R to test whether there is evidence that Doris and Buzz can communicate.



$\\$






**Part 1.1 (5 points)**:  Step 1 of hypothesis testing: state the null and alternative hypothesis. Also state significance level $\alpha$ level.









$\\$






**Part 1.2 (3 points) **: Step 2 of hypothesis testing: calculate the statistic of interest. 




```{r problem_1_2, message=FALSE, warning=FALSE, tidy=TRUE}


  
  # calculate the observed statistic

  
    
```






$\\$






**Part 1.3 (5 points) **: Step 3 of hypothesis testing: create the null distribution. 



```{r message=FALSE, warning=FALSE, tidy=TRUE}
    



        
```





$\\$





**Part 1.4 (10 points)**: Step 4 of hypothesis testing: calculate the p-value. 




```{r problem_1_4, message=FALSE, warning=FALSE, tidy=TRUE}
    






    
```







$\\$





**Part 1.5 (5 points)** Step 5 of hypothesis testing: make a judgment. 







