---
title: "Class 7 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```






$\\$







## Part 1: Parametric tests for comparing two means



**Question:** Does eating late at night make you put on more weight compared to whether you ate the same amount of food during the day? 


To test this idea researchers, [Fonken et al, 2010](https://www.ncbi.nlm.nih.gov/pubmed/20937863), examined this question in mice. 

Mice were randomly divided into 2 groups:

 * **Dark condition**:  8 mice were given 8 hours of darkness at night when they couldn’t eat
 * **Light condition**: 9 were constantly exposed to light for 24 hours so they could always eat


The data from the mice is below. Let's use a t-test to see if there is any difference between the amount of weight the mice gained between these two conditions. 


Our formula for a t-statistic for comparing two means is:  $t = \frac{\bar{x_{light}} - \bar{x_{dark}}}{SE}$ 


Our formula for calculating the standard error is: $SE = \sqrt{\frac{s_1^2}{n_2} + \frac{s_2^2}{n_1}}$


This is Welch's t-test, which gives valid results even when there are unequal variances between the two groups. 





$\\$





##### Step 0: Plot the data


```{r mice_plot}

# run this once
# get the mice data
# source('class_functions.R')
# download_class_data("mice.Rda")

# load the data
load("mice.Rda")




# a) creating a list of the data 



# b) creating a stripchart   



```




##### Step 1: State the null and alternative hypotheses 








$\\$







##### Step 2: Calculate the observed statistic

```{r pill_obs_stat}






```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_mice}



# sample sizes for the two conditions



# a conservative estimate of the degrees of freedom




# plot the null distribution with the observed statistic on it









```






$\\$





##### Step 4: Calculate a p-value


```{r p_value_mice}




# calculate the p-value






```




$\\$






##### Step 5: Make a decision









$\\$








## Part 2: Using the bootstrap to create confidence intervals


Let's create a confidence interval for the mean amount of weight ($\mu_{light}$) that we would expect our mice in the light group to put on using the bootstrap. 




$\\$



**Part 2.1**  Use a for loop to create a bootstrap distribution by: 

1) sampling with replacement *n* values from light weight gain group
2) calculating the mean of the bootstrap sample and storing it in a vector named `bootstrap_dist`
3) repeating this process 10,000 times. 

Then plot a histogram of the bootstrap distribution and calculate the bootstrap standard error $SE*$ and store it in an object called `SE_boot`

```{r message=FALSE, warning=FALSE, tidy=TRUE}


# get the sample size



# run a for loop to generate the bootstrap distribution 








```



$\\$






**Part 2.2** 

Let's use the bootstrap standard error you to now calculate a 95% confidence interval using the formula:  

$CI_b = \bar{x} \pm  2 \cdot SE*$. 


We can also use the quantile() function to get a 95% confidence interval.



```{r}


# get the observed statistic


# use the formula to get the bootstrap confidence interval


# also get the CIs using quantiles...


```








$\\$







**Part 2.3**  We can also calculate confidence intervals for the mean $\mu$ using the formula:

$CI = \bar{x} \pm 2 \cdot SE$

where SE is given by the formula $SE = \frac{s}{\sqrt{n}}$ and  *s* is the standard deviation of the sample and *n* is the sample size. 


```{r}

# calculate the standard error




# create the confidence interval




```





$\\$




**Part 2.4** The value of 2 in the equation for the the confidence interval $CI = \bar{x} \pm 2 \cdot SE$ is an approximate value. We can get a more precise confidence interval if we use the quantile value from a t-distribution with n-1 degrees of freedom, i.e. we should use the formula $CI = \bar{x} \pm t^* \cdot SE$. 

We can get such a quantile value $t^*$ from a t-distribution in R using the `qt(p, df)` function:



```{r}


# get the quantile value from a t-distribution


# calculate the CI using this quantile value




```



