---
title: "Class 5 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview

 * Binomial density functions and binomial cumulative distributions functions
 * Normal density functions and normal cumulative distributions functions
 * For loops
 


$\\$



## Part 1: Hypothesis tests for a single proportion using the binomial distribution


Let us use R to test whether there is evidence that Doris and Buzz can communicate. This time we will use a **binomial density function** as the null distribution rather than just generated samples from a binomial distribution. 


Recall, for step 1 of hypothesis testing, the null and alternative hypotheses for this problem were: 


$H_0: \pi = 0.5$ 
$H_A: \pi > 0.5$ 



Also, recall for step 2 of hypothesis testing, the observed statistic was:  $\hat{p} = 15/16$


```{r}





```





$\\$



**Part 1.1**:  Let's create a binomial null distribution consistent with the null hypothesis (step 3 of hypothesis testing). Here we will plot the full **density function**. 


```{r}

# use the dbinom() function to create a probability density for all values from 0 to 16



# add names to this vector



# use the barplot to plot the density function 




```



$\\$




**Part 1.2**:  Let's now calculate a p-value from this null distribution and the observed statistic (step 4 of hypothesis testing). To do this we will use the **cumulative distribution function** to get $Pr(K \ge obsstat; n, \pi)$


```{r}

# use the cumulative distribution function pbinom() to get the p-value  Pr(k > 14)


# verify that it is the same as adding the tail of the density function 



```


Now based on this p-value, do step 5 of hypothesis testing and make a decision 






$\\$







## Part 2: Hypothesis tests for a single proportion using the normal distribution


We can also use the normal density function as an approximation for the binomial density function when calculating p-values. Again, steps 1 and 2 of hypothesis testing are the same: 


Step 1:

$H_0: \pi = 0.5$ 
$H_A: \pi > 0.5$ 


Step 2: 

$\hat{p} = 15/16$






$\\$







**Part 2.1**:  Let's create a normal null distribution consistent with the null hypothesis (step 3 of hypothesis testing). Here we will plot the full **density function**. 


Recall, when using a normal distribution as an approximation for the binomal distribution we have the follow values for the parameters of the normal distribution 


$\mu = \pi_0$

$\sigma = \sqrt{\frac{\pi (1 - \pi)}{n}}$



```{r}


# calculate the values fo mu and sigma



# for plotting the density function, get a range of x-values




# get the corresponding normal density y-values





# plot the density function 





```




$\\$



**Part 2.2**:  Let's now calculate a p-value from this null distribution and the observed statistic (step 4 of hypothesis testing). Do do this we will use the **cumulative distribution function** to get $Pr(K \ge obsstat; n, \pi)$


```{r}

# use the pnorm function to get the p-value Pr(X > obs_stat | H0 parameters)



```


Now based on this p-value, do step 5 of hypothesis testing and make a decision




$\\$






## Part 3: R coding - for loops

For loops are useful when you want to repeat a piece of code many times under similar conditions.



**Part 3.1**: In every iteration of a for loop, the value of *i* is increated by 1.



```{r}




```


$\\$




**Part 3.2**:  For loops are particular useful in combination with vectors that can store the results


```{r}


# create an empty vector to store the results


# add values to this vector in a for loop




```


$\\$














