---
title: "Class 3 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview

 * Learning about R Markdown
 * Back to the basics of R
 * Generated random numbers from a binomial distribution
 




$\\$





#### Part 0: R Markdown


R Markdown allows you to combine a written report along with code. All homework assignments will be in R Markdown. 


You can also embed links to websites and special $LaTeX$ characters like $\pi$ and $\bar{x}$ and you can put works in **bold** and *italics*








$\\$






#### Part 1.1 Assigning values to objects


As we saw last class, we can assign numbers to objects



```{r}

a <- 7
b <- 52
z  <- a * b
z


```




$\\$




#### Part 2.1: Numeric, string and boolean data types

We can also assign strings and booleans to objects. We can get the "type" of an object using the class() function. 



```{r}

a <- 7			
s <- "s is a terrible name for an object"
b <- TRUE

class(a)
class(s)


```




$\\$




#### Part 3.1: Functions

We can use function to do computations and return results



```{r}


sqrt(49)			
tolower("DATA is AWESOME!")


```



$\\$




#### Part 3.2: Getting help

To get help information about a function or data set use ?  


How else can you get help on how to use R?

[Help desk cat](https://tr4.cbsistatic.com/hub/i/2017/06/12/bd6094f4-c445-4619-a2a4-376b3a8c2065/google-business-cat.jpg)

[Another way to get help](https://piazza.com/class/k2cifwdm13c1kv)




```{r}

? sqrt


```



$\\$




#### Part 4.1: One can add comments to your code to describe what you are doing

```{r}

sqrt(49)    # this takes the square root of 49


```



$\\$




#### Part 5.1: Vectors are ordered sequences of numbers or letters


The c() function is used to create vectors


```{r}


v  <-  c(5, 232, 5, 543)	


```



$\\$





#### Part 5.2: One can access elements of a vector using square brackets []


```{r}

v[4]         # what will the answer be?    


```




$\\$




#### Part 5.3: We can get multiple elements from a vector too

```{r}

v[c(1, 2)]


```



$\\$




#### Part 5.4: One can assign a sequence of numbers to a vector


```{r}

z <- 2:10
z[3]	


```



$\\$




#### Part 5.5: One can test which elements are greater than a value

```{r}

z > 3
sum(z > 3)


```



$\\$




#### Part 5.6: One can also apply functions to vectors

```{r}

z <- 2:10
sqrt(z)
mean(z)

```




$\\$





#### Part 5.7: We can create a vector of booleans for elements are greater than a value

```{r}

the_trues <- z > 3


```




$\\$




#### Part 5.8: TRUEs are treated as 1’s and FALSEs are treated as 0


```{r}

sum(the_trues)     # count how many elements are TRUE


```



$\\$




#### Part 5.9: We can use rbinom() to generate random numbers from a binomial distribution

The rbinom() function has the following arguments

rbinom(num_sims, size, prob)

 * num_sims: the number of simulations (random numbers) we want to generate
 * size:  the number of coin flips on each simulation  (i.e., n)
 * prob: the probability of a “head” on each coin flip (i.e., p or $\pi$)


Let's 100 random numbers k, from $P(k; n = 20, \pi = .75)$

```{r}

flip_simulations <- rbinom(100, 20, .75)


```



$\\$




#### Part 5.10: We can count how many times we got k heads using:

```{r}

count_num_heads <- table(flip_simulations)


```



$\\$




#### Part 5.11: We can plot this as a barplot using


```{r}

barplot(count_num_heads, 
        xlab = "Number of heads", 
        ylab = "Count")


```



$\\$



You are a [rockstar](https://images2.minutemediacdn.com/image/upload/c_fill,g_auto,h_1248,w_2220/f_auto,q_auto,w_1100/v1555319921/shape/mentalfloss/matrix_header.jpg)










