---
title: "Class 8 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```






$\\$










## Part 1: Using the bootstrap to create confidence intervals


Let's create a confidence interval for the mean amount of weight ($\mu_{light}$) that we would expect our mice in the light group to put on using the bootstrap. 



```{r}

# download_class_data("mice.Rda")

# load the data
load("mice.Rda")


```




$\\$



**Part 1.1**  Use a for loop to create a bootstrap distribution by: 

1) sampling with replacement *n* values from light weight gain group
2) calculating the mean of the bootstrap sample and storing it in a vector named `bootstrap_dist`
3) repeating this process 10,000 times. 

Then plot a histogram of the bootstrap distribution and calculate the bootstrap standard error $SE*$ and store it in an object called `SE_boot`

```{r message=FALSE, warning=FALSE, tidy=TRUE}


# get the sample size
n_sample_size <- length(light_BM_increase)


# run a for loop to generate the bootstrap distribution 
bootstrap_dist <- NULL
for (i in 1:10000) {    
  
  curr_boot <- sample(light_BM_increase, n_sample_size, replace = TRUE)  
  bootstrap_dist[i] <-  mean(curr_boot)  
  
}


# create a histogram of the bootstrap distribution
hist(bootstrap_dist, nclass = 100, main = "Bootstrap distribution of mean weight gains", xlab = "Mice mean weight gains x-bar")


# calculate the bootstrap standard error
SE_boot <- sd(bootstrap_dist)
SE_boot




```



$\\$






**Part 1.2** 

Let's use the bootstrap standard error you to now calculate a 95% confidence interval using the formula:  

$CI_b = \bar{x} \pm  2 \cdot SE*$. 


We can also use the quantile() function to get a 95% confidence interval.



```{r}


the_stat <- mean(light_BM_increase)


# use the formula to get the bootstrap confidence interval
CI <- the_stat + 2 * SE_boot * c(-1, 1)
CI


# also get the CIs using quantiles...
quantile(bootstrap_dist, c(.025, .975))


```








$\\$







**Part 1.3**  We can also calculate confidence intervals for the mean $\mu$ using the formula:

$CI = \bar{x} \pm 2 \cdot SE$

where SE is given by the formula $SE = \frac{s}{\sqrt{n}}$ and  *s* is the standard deviation of the sample and *n* is the sample size. 


```{r}

# calculate the standard error
the_sd <- sd(light_BM_increase)
the_se <- the_sd/sqrt(n_sample_size)

# create the confidence interval
(CI_formula <- the_stat + c(-2, 2) * the_se)




```





$\\$




**Part 1.4** The value of 2 in the equation for the the confidence interval $CI = \bar{x} \pm 2 \cdot SE$ is an approximate value. We can get a more precise confidence interval if we use the quantile value from a t-distribution with n-1 degrees of freedom, i.e. we should use the formula $CI = \bar{x} \pm t^* \cdot SE$. 

We can get such a quantile value $t^*$ from a t-distribution in R using the `qt(p, df)` function:



```{r}


# get the quantile value from a t-distribution
(t_star <- qt(.975, df = n_sample_size - 1))


# calculate the CI using this quantile value
(CI_formula2 <- the_stat + c(-t_star, t_star) * the_se)


```



