---
title: "Class 9 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)


```







```{r}



# some helper functions


# a function to get the MAD statistic
get_MAD_stat <- function(data, grouping) {
  
  group_means <-  as.vector(by(data, grouping, mean))
  
  total <- 0
  num_differences <- 0
  
  for (iGroup1 in 1:(length(group_means) - 1)) {
  
    for (iGroup2 in (iGroup1 + 1):(length(group_means))) {
    
      total <- total + abs(group_means[iGroup1] - group_means[iGroup2])
      num_differences <- num_differences + 1
    
    }
  
  }

  
  total/num_differences

}  # end of the function



# A function to get the F-statistic
get_F_statistic <- function(data, grouping) {
  fit <- aov(data ~ grouping)
  fit_summary <- summary.aov(fit)
  fit_summary[[1]]$`F value`[1]
}




```





$\\$





## Part 1: Permutation tests for testing more than two means 


Mercury poisoning can lead to a range of physical and cognitive symptoms. Several studies have examined the amount of mercury in fish which is a source of mercury poisoning in humans - for example, [see this paper](https://ecos.fws.gov/ServCat/DownloadFile/21887). 

We will examine a data set that measured the amount of mercury in 4 species of fish to see if there is a difference in the mean mercury levels between species. 





```{r}

mercury <- read.csv("mercury.csv")


```




$\\$






**Part 1.0 - step 0** Let's start our analysis by describing and plotting the data by:

1) Counting how many cases are in the data frame, and what each case corresponds to
2) Creating a side-by-side boxplot comparing the mercury levels in these different species


Does it appear that the mercury level differs on average depending on the species of fish? 




```{r visualize_fish}

# calculate how many fish in total
dim(mercury)


# calculate how many of each species 
table(mercury$Fish)


# create a boxplot comparing the mercury levels in the different species of fish
boxplot(MercuryLevel ~ Fish, data = mercury, 
        xlab = "Species",
        ylab = "Mercury level (ppm)")



```





$\\$







**Part 1.1 - step 1**:  Let's examine whether there is a statistically significant difference in the mean mercury level for between the different species. Start by stating the null and alternative hypotheses in symbols and words, and also state the alpha level that is most commonly used. 




**In words**

The null hypothesis is that there is no difference between the mean mercury level between different species. The alternative hypothesis is that some species have different mean mercury levels. 


**In symbols**

$H_0: \mu_B = \mu_H = \mu_O = \mu_P$
$H_A: \mu_i \ne \mu_j$ for some i = j, where i, j are in {Bass, Halibut, Orange-roughy and Perch}



**The significance level**

The significance level is probability of falsely rejected $H_0$ when it is true. The typical level used is $\alpha = 0.05$




$\\$





**Part 1.2 - step 2**: Calculate the observed statistic.

Let's now compute of statistic of interest. There are many statistics we could use, but let's use the MAD statistic. I have created a function (above) called `get_MAD_stat()` which takes a vector of quantitative data, and a vector specifying the group membership and will return the MAD statistic. Let's use it to calculate the MAD statistic for the mercury level in the different species of fish. 


```{r}

(obs_stat <- get_MAD_stat(mercury$MercuryLevel, mercury$Fish))

# (obs_stat <- get_F_statistic(mercury$MercuryLevel, mercury$Fish))


```



$\\$





**Part 1.3 - step 3**: Create the null distribution. 


We can create the null distribution by:

1) shuffling which mercury level belongs to which fish
2) calculating the mercury level for the fake fish groups
3) repeating this process 10,000 times



```{r}


null_distribution <- NULL
for (i in 1:10000){
  
  null_distribution[i] <- get_MAD_stat(mercury$MercuryLevel, sample(mercury$Fish))
  # null_distribution[i] <- get_F_statistic(mercury$MercuryLevel, sample(mercury$Fish))
  
}


hist(null_distribution, breaks = 100)




```






$\\$





**Part 1.4 - step 4**: Calculate the p-value.

Let's calculate the p-value by seeing how many points in our null distribution are as great or greater than the observed statistic.


```{r}


(p_val <- sum(null_distribution >= obs_stat))



```



$\\$





**Part 1.5 - step 5**: Make a decision!



[Fish food](https://i.etsystatic.com/11478882/r/il/fd4e11/1075438060/il_794xN.1075438060_3ym5.jpg)





$\\$



