---
title: "Class 9 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)


```







```{r}



# some helper functions


# a function to get the MAD statistic
get_MAD_stat <- function(data, grouping) {
  
  group_means <-  as.vector(by(data, grouping, mean))
  
  total <- 0
  num_differences <- 0
  
  for (iGroup1 in 1:(length(group_means) - 1)) {
  
    for (iGroup2 in (iGroup1 + 1):(length(group_means))) {
    
      total <- total + abs(group_means[iGroup1] - group_means[iGroup2])
      num_differences <- num_differences + 1
    
    }
  
  }

  
  total/num_differences

}  # end of the function



# A function to get the F-statistic
get_F_statistic <- function(data, grouping) {
  fit <- aov(data ~ grouping)
  fit_summary <- summary.aov(fit)
  fit_summary[[1]]$`F value`[1]
}




```





$\\$





## Part 1: Permutation tests for testing more than two means 


Mercury poisoning can lead to a range of physical and cognitive symptoms. Several studies have examined the amount of mercury in fish which is a source of mercury poisoning in humans - for example, [see this paper](https://ecos.fws.gov/ServCat/DownloadFile/21887). 

We will examine a data set that measured the amount of mercury in 4 species of fish to see if there is a difference in the mean mercury levels between species. 





```{r}


# download.file('https://raw.githubusercontent.com/emeyers/SDS101_F19/master/class_data/mercury.csv', 'mercury.csv')

# mercury <- read.csv("mercury.csv")


```




$\\$






**Part 1.0 - step 0** Let's start our analysis by describing and plotting the data by:

1) Counting how many cases are in the data frame, and what each case corresponds to
2) Creating a side-by-side boxplot comparing the mercury levels in these different species


Does it appear that the mercury level differs on average depending on the species of fish? 




```{r visualize_fish}


# calculate how many fish in total



# calculate how many of each species 



# create a boxplot comparing the mercury levels in the different species of fish




```





$\\$







**Part 1.1 - step 1**:  Let's examine whether there is a statistically significant difference in the mean mercury level for between the different species. Start by stating the null and alternative hypotheses in symbols and words, and also state the alpha level that is most commonly used. 




**In words**

The null hypothesis is that there is no difference between the mean mercury level between different species. The alternative hypothesis is that some species have different mean mercury levels. 


**In symbols**




**The significance level**






$\\$





**Part 1.2 - step 2**: Calculate the observed statistic.

Let's now compute of statistic of interest. There are many statistics we could use, but let's use the MAD statistic. I have created a function (above) called `get_MAD_stat()` which takes a vector of quantitative data, and a vector specifying the group membership and will return the MAD statistic. Let's use it to calculate the MAD statistic for the mercury level in the different species of fish. 


```{r}




```



$\\$





**Part 1.3 - step 3**: Create the null distribution. 


We can create the null distribution by:

1) shuffling which mercury level belongs to which fish
2) calculating the mercury level for the fake fish groups
3) repeating this process 10,000 times



```{r}







```






$\\$





**Part 1.4 - step 4**: Calculate the p-value.

Let's calculate the p-value by seeing how many points in our null distribution are as great or greater than the observed statistic.


```{r}





```



$\\$





**Part 1.5 - step 5**: Make a decision!



[Fish food](https://i.etsystatic.com/11478882/r/il/fd4e11/1075438060/il_794xN.1075438060_3ym5.jpg)





$\\$





## Part 2: One-way analysis of variance (ANOVA) for testing more than two means 


We can also run an analysis of variance (ANOVA) for testing if there are differences between more than 2 means. Let's continue to examine the mercury in fish data. 


To start this analysis we state the same null and alternative hypothesis as when doing a permutation test, so let's go straight to step 2. 




$\\$




**Part 1.2 - steps 2 to 5**:  A ANOVA is a parametric hypothesis test based on using an F-statistic. The equation for an F-statistic is:

$$F ~ = ~ \frac{ \mbox{between-group variability}}{\mbox{within-group variability}} ~ = ~ \frac{\frac{1}{K - 1}\sum_{i=1}^{K}{n_i(\bar{x}_i - \bar{x}_{tot})^2}}{\frac{1}{N - K}\sum_{i=1}^{K}\sum_{j=1}^{n_i}{(x_{ij}-\bar{x}_i)^2}}$$

In the above equation, the symbols mean the following:

 * K = 4 corresponds to the 4 species levels (Bass, Halibut, Orange-roughy and Perch)
 * N corresponds to the total number of fish we are using in our analysis
 * $x_{ij}$ corresponds to the $j^{th}$ mercury level in the $i^{th}$ species of fish
 * $n_i$ is the number of fish in the $i^{th}$ group (e.g., the number of Bass)
 * $\bar{x}_i$ is the average score for the $i^{th}$ group (e.g., the average mercury level for Bass)


Writting the code out to calculate the F-statistic seems like a lot of work. Let's cheat and just use R's built in ANOVA functions... 



```{r F_permutation_test}




```


How does R's ANOVA function work? Conceptually you can think of it as computing the F-statistic and using the F-distribution to get the p-value, although R actually solves this problem in another equivalent way using code written to do regression which gives the same answer. 






$\\$




**Part 1.3 - checking ANOVA assumptions**: 


The results from an ANOVA are valid if a set of assumptions are met. These assumptions are: 






Let's check them now.



```{r}

# visualize data in each group (could also do qqplot to examine normality)



# not very normal :(  - although ANOVAs are pretty robust to this



# standard deviations fairly similar in each group



```


The data in the groups is not normally distributed, however the ANOVA is often very robust to a violation of the normal assumption [as can be seen in this paper by Blanca et al 2017](https://www.ncbi.nlm.nih.gov/pubmed/29048317)





$\\$
