---
title: "Class 10 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)


```







```{r}



# some helper functions


# a function to get the MAD statistic
get_MAD_stat <- function(data, grouping) {
  
  group_means <-  as.vector(by(data, grouping, mean))
  
  total <- 0
  num_differences <- 0
  
  for (iGroup1 in 1:(length(group_means) - 1)) {
  
    for (iGroup2 in (iGroup1 + 1):(length(group_means))) {
    
      total <- total + abs(group_means[iGroup1] - group_means[iGroup2])
      num_differences <- num_differences + 1
    
    }
  
  }

  
  total/num_differences

}  # end of the function



# A function to get the F-statistic
get_F_statistic <- function(data, grouping) {
  fit <- aov(data ~ grouping)
  fit_summary <- summary.aov(fit)
  fit_summary[[1]]$`F value`[1]
}




```




$\\$





## Part 1: One-way analysis of variance (ANOVA) for testing more than two means 


We run an analysis of variance (ANOVA) for testing if there are differences between more than 2 means. Let's continue to examine the mercury in fish data. 





$\\$



```{r}


# download.file('https://raw.githubusercontent.com/emeyers/SDS101_F19/master/class_data/mercury.csv', 'mercury.csv')


mercury <- read.csv("mercury.csv")



```






$\\$






**Part 1.0 - step 0** Let's start our analysis by creating a strip chart to view our data



```{r visualize_fish}



# create a boxplot comparing the mercury levels in the different species of fish
# boxplot(MercuryLevel ~ Fish, data = mercury, 
#        xlab = "Species",
#        ylab = "Mercury level (ppm)")




# create a stripchart comparing the mercury levels in the different species of fish
stripchart(MercuryLevel ~ Fish, data = mercury, 
           method = "jitter", 
           vertical = TRUE, 
           pch = 1,
           xlab = "Species",
           ylab = "Mercury level (ppm)")


```





$\\$







**Part 1.1 - step 1**:  Let's examine whether there is a statistically significant difference in the mean mercury level for between the different species. Start by stating the null and alternative hypotheses in symbols and words, and also state the alpha level that is most commonly used. 




**In words**

The null hypothesis is that there is no difference between the mean mercury level between different species. The alternative hypothesis is that some species have different mean mercury levels. 


**In symbols**

$H_0: \mu_B = \mu_H = \mu_O = \mu_P$
$H_A: \mu_i \ne \mu_j$ for some i = j, where i, j are in {Bass, Halibut, Orange-roughy and Perch}



**The significance level**

The significance level is probability of falsely rejected $H_0$ when it is true. The typical level used is $\alpha = 0.05$




$\\$








**Part 1.2 - steps 2 to 5**:  A ANOVA is a parametric hypothesis test based on using an F-statistic. The equation for an F-statistic is:

$$F ~ = ~ \frac{ \mbox{between-group variability}}{\mbox{within-group variability}} ~ = ~ \frac{\frac{1}{K - 1}\sum_{i=1}^{K}{n_i(\bar{x}_i - \bar{x}_{tot})^2}}{\frac{1}{N - K}\sum_{i=1}^{K}\sum_{j=1}^{n_i}{(x_{ij}-\bar{x}_i)^2}}$$

In the above equation, the symbols mean the following:

 * K = 4 corresponds to the 4 species levels (Bass, Halibut, Orange-roughy and Perch)
 * N corresponds to the total number of fish we are using in our analysis
 * $x_{ij}$ corresponds to the $j^{th}$ mercury level in the $i^{th}$ species of fish
 * $n_i$ is the number of fish in the $i^{th}$ group (e.g., the number of Bass)
 * $\bar{x}_i$ is the average score for the $i^{th}$ group (e.g., the average mercury level for Bass)


Writting the code out to calculate the F-statistic seems like a lot of work. Let's cheat and just use R's built in ANOVA functions... 



```{r F_permutation_test}

fit <- aov(MercuryLevel ~ Fish, data = mercury)

summary.aov(fit)


```


How does R's ANOVA function work? Conceptually you can think of it as computing the F-statistic and using the F-distribution to get the p-value, although R actually solves this problem in another equivalent way using code written to do regression which gives the same answer. 






$\\$




**Part 1.3 - checking ANOVA assumptions**: 


The results from an ANOVA are valid if a set of assumptions are met. These assumptions are: 

 * Data in each group come from normal distributions 
 * Each group has equal variance  (homoskedasticity) 



Let's check these assumptions now...



```{r}

# visualize data in each group (could also do qqplot to examine normality)
#stripchart(MercuryLevel ~ Fish, 
#           data = mercury, 
#           method = "jitter", 
#           vertical = TRUE, 
#           pch = 1,
#           ylab = "Mercurcy level (ppm)", 
#           xlab = "Species")


library(ggplot2)
qplot(Fish, MercuryLevel, data = mercury, geom = "violin")


# not very normal :(  - although ANOVAs are pretty robust to this
bass <- subset(mercury, mercury$Fish == "bass")
qqnorm(bass$MercuryLevel)
hist(bass$MercuryLevel)


# standard deviations fairly similar in each group
by(mercury$MercuryLevel, mercury$Fish, sd)


```


The data in the groups is not normally distributed, however the ANOVA is often very robust to a violation of the normal assumption [as can be seen in this paper by Blanca et al 2017](https://www.ncbi.nlm.nih.gov/pubmed/29048317)





$\\$
