---
title: "Class 6 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$







## Part 1: R coding - for loops

For loops are useful when you want to repeat a piece of code many times under similar conditions.



**Part 1.1**: In every iteration of a for loop, the value of *i* is increated by 1.



```{r}




```


$\\$




**Part 1.2**:  For loops are particular useful in combination with vectors that can store the results


```{r}




```






$\\$







## Part 2: Permutation tests for comparing two sample means in R



##### Step 0: Plot the data


```{r pill_plot}



treat <- c(7, -4, 18, 17, -3, -5,  1, 10, 11, -2)
control <- c(-1,  12,  -1,  -3,   3,  -5,   5,   2, -11,  -1,  -3)


# alternatively, could do the mice gaining weight example
# download_class_data('mice.Rda')
# dark_BM_increase <- c(9.58,  7.69,  9.83,  8.70, 13.29,  7.24,  9.72,  5.58)
# light_BM_increase <- c(11.98,  9.97,  9.79, 10.24, 10.28,  7.97, 18.91, 12.14, 15.32)


# plot the data using a boxplot





```







##### Step 1: State the null and alternative hypotheses 







$\\$




##### Step 2: Calculate the observed statistic

```{r pill_obs_stat}





```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_pill}


# combine the data from the treatment and control groups together





# use a for loop to create shuffled treatment and control groups and shuffled statistics 





# plot the null distribution as a histogram




```



$\\$




##### Step 4: Calculate a p-value


```{r p_value_pill}


# plot the null distribution again with a red line a the value of the observed statistic






# calculate the p-value







```




$\\$



##### Step 5: Make a decision






$\\$





[pills](https://raw.githubusercontent.com/emeyers/SDS230_F19/master/class_images/class_09/pills.jpg)






$\\$







## Part 3: Permutation tests for comparing two sample means in R - mice example



**Question:** Does eating late at night make you put on more weight compared to whether you ate the same amount of food during the day? 


To test this idea researchers, [Fonken et al, 2010](https://www.ncbi.nlm.nih.gov/pubmed/20937863), examined this question in mice. 

Mice were randomly divided into 2 groups:

 * **Dark condition**:  8 mice were given 8 hours of darkness at night when they couldn’t eat
 * **Light condition**: 9 were constantly exposed to light for 24 hours so they could always eat


The data from the mice is below. Let's use a permutation test to see if there is any difference between the amount of weight the mice gained between these two conditions. 




$\\$




##### Step 0: Plot the data


```{r mice_plot}

# run this once
# get the mice data
# source('class_functions.R')
# download_class_data("mice.Rda")

# load the data
load("mice.Rda")



# plot the data as a boxplot





```




$\\$




##### Step 1: State the null and alternative hypotheses 







$\\$





##### Step 2: Calculate the observed statistic

```{r mice_obs_stat}




```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_mice}


# combine the data from the dark and light groups together




# use a for loop to create shuffled dark and light groups and shuffled statistics 






```





$\\$





##### Step 4: Calculate a p-value


```{r p_value_mice}


# plot the null distribution again with a red line a the value of the observed statistic



# calculate the p-value




```




$\\$






##### Step 5: Make a decision




$\\$





[Give a mouse a cookie](https://hamstersworld.weebly.com/uploads/2/0/6/1/2061190/8024088.jpg)


[science has gone too far](https://i.imgflip.com/9vr.jpg)



