---
title: "Class 5 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview

 * Binomial density functions and binomial cumulative distributions functions
 * Normal density functions and normal cumulative distributions functions




$\\$



## Part 1: Hypothesis tests for a single proportion using the binomial distribution


Let us use R to test whether there is evidence that Doris and Buzz can communicate. This time we will use a **binomial density function** as the null distribution rather than just generated samples from a binomial distribution. 


Recall, for step 1 of hypothesis testing, the null and alternative hypotheses for this problem were: 


$H_0: \pi = 0.5$ 
$H_A: \pi > 0.5$ 



Also, recall for step 2 of hypothesis testing, the observed statistic was:  $\hat{p} = 15/16$


```{r}

obs_stat <- 15/16


```





$\\$



**Part 1.1**:  Let's create a binomial null distribution consistent with the null hypothesis (step 3 of hypothesis testing). Here we will plot the full **density function**. 


```{r}


null_distribution  <-  dbinom(0:16, 16, .5)/16
names(null_distribution) <- (0:16)/16

  
barplot(null_distribution, xlab = "p-hat", ylab = "Relative frequency")



```



$\\$




**Part 1.2**:  Let's now calculate a p-value from this null distribution and the observed statistic (step 4 of hypothesis testing). Do do this we will use the **cumulative distribution function** to get $Pr(K \ge obsstat; n, \pi)$


```{r}


(p_value  <-  pbinom(14, 16, .5, lower.tail = FALSE))  # lower.tail = FALSE -> Pr(k > 14)
                                                       # lower.tail = TRUE -> Pr(k < 14)

dbinom(15, 16, .5) + dbinom(16, 16, .5)


```


Now based on this p-value, do step 5 of hypothesis testing and make a decision 






$\\$







## Part 2: Hypothesis tests for a single proportion using the normal distribution


We can also use the normal density function as an approximation for the binomial density function when calculating p-values. Again, steps 1 and 2 of hypothesis testing are the same: 


Step 1:

$H_0: \pi = 0.5$ 
$H_A: \pi > 0.5$ 


Step 2: 

$\hat{p} = 15/16$






$\\$







**Part 2.1**:  Let's create a normal null distribution consistent with the null hypothesis (step 3 of hypothesis testing). Here we will plot the full **density function**. 


Recall, when using a normal distribution as an approximation for the binomal distribution we have the follow values for the parameters of the normal distribution 


$\mu = \pi_0$

$\sigma = \sqrt{\frac{\pi (1 - \pi)}{n}}$



```{r}


mu <- .5
sigma <- sqrt(  (.5 * (1 - .5)) / 16)


x_range <- seq(-.5, 1.5, by = .01)

y_density_vals  <-  dnorm(x_range, mu, sigma)

  
plot(x_range, y_density_vals, type = "l", 
     ylab = "Relative frequency", 
     xlab = "p-hat approximation")


```




$\\$



**Part 2.2**:  Let's now calculate a p-value from this null distribution and the observed statistic (step 4 of hypothesis testing). Do do this we will use the **cumulative distribution function** to get $Pr(K \ge obsstat; n, \pi)$


```{r}


(p_value  <-  pnorm(15/16, mu, sigma, lower.tail = FALSE))



```


Now based on this p-value, do step 5 of hypothesis testing and make a decision




$\\$










