---
title: "Class 6 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$







## Part 1: R coding - for loops

For loops are useful when you want to repeat a piece of code many times under similar conditions.



**Part 1.1**: In every iteration of a for loop, the value of *i* is increated by 1.



```{r}

for (i in 1:20) {
	print(i)
}


```


$\\$




**Part 1.2**:  For loops are particular useful in combination with vectors that can store the results


```{r}


my_results <- NULL      # create an empty vector to store the results
for (i in 1:20) {
	 my_results[i] <- i^2
}


my_results

```






$\\$







## Part 2: Permutation tests for comparing two sample means in R



##### Step 0: Plot the data


```{r pill_plot}



treat <- c(7, -4, 18, 17, -3, -5,  1, 10, 11, -2)
control <- c(-1,  12,  -1,  -3,   3,  -5,   5,   2, -11,  -1,  -3)



# plot the data
boxplot(treat, control, 
        names = c("Treatment", "Control"),
        main = "Comparison of decrease in blood pressure",
        ylab = "Blood pressure decrease")



# create a stripchart
data_list <- list(treat, control) 
  
stripchart(data_list, 
           group.names = c("treatment", "control"), 
           main = "Comparison of decrease in blood pressure",
           xlab = "Blood pressure decrease", 
           col = c("red", "blue"))


```







##### Step 1: State the null and alternative hypotheses 


$H_0: \mu_{treat} - \mu_{control} = 0$
$H_A: \mu_{treat} - \mu_{control} > 0$


$\alpha = 0.05$




$\\$




##### Step 2: Calculate the observed statistic

```{r pill_obs_stat}


obs_stat <- mean(treat) - mean(control)



```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_pill}


# combine the data from the treatment and control groups together
combined_data <- c(treat, control)


# use a for loop to create shuffled treatment and control groups and shuffled statistics 
null_distribution <-  NULL
for (i in 1:10000) {

       # shuffle data
        shuff_data <- sample(combined_data)

        # create fake treatment and control groups
        shuff_treat   <-  shuff_data[1:10]
        shuff_control  <-  shuff_data[11:21]

        # save the statistic of interest
        null_distribution[i] <- mean(shuff_treat) - mean(shuff_control)

        
}


# plot the null distribution as a histogram
hist(null_distribution, 
     nclass = 200,
     main = "Null distribution", 
     xlab = TeX("$\\bar{x}_{shuff-treat} - \\bar{x}_{shuff-control}$"))




```



$\\$




##### Step 4: Calculate a p-value


```{r p_value_pill}


# plot the null distribution again with a red line a the value of the observed statistic
hist(null_distribution, 
     nclass = 200,
     main = "Null distribution", 
     xlab = TeX("$\\bar{x}_{shuff-treat} - \\bar{x}_{shuff-control}$"))


abline(v = obs_stat, col = "red")



# calculate the p-value
(p_value <- sum(null_distribution >= obs_stat)/10000)




```




$\\$



##### Step 5: Make a decision

Since `r p_value` is greater than $\alpha = 0.05$ we can not reject the null hypothesis. Thus if we are using the Neyman-Pearson paradigm, we do not have sufficient evidence to say that the pill is effective. 



$\\$





[pills](https://raw.githubusercontent.com/emeyers/SDS230_F19/master/class_images/class_09/pills.jpg)






$\\$







## Part 3: Permutation tests for comparing two sample means in R - mice example



**Question:** Does eating late at night make you put on more weight compared to whether you ate the same amount of food during the day? 


To test this idea researchers, [Fonken et al, 2010](https://www.ncbi.nlm.nih.gov/pubmed/20937863), examined this question in mice. 

Mice were randomly divided into 2 groups:

 * **Dark condition**:  8 mice were given 8 hours of darkness at night when they couldn’t eat
 * **Light condition**: 9 were constantly exposed to light for 24 hours so they could always eat


The data from the mice is below. Let's use a permutation test to see if there is any difference between the amount of weight the mice gained between these two conditions. 




$\\$




##### Step 0: Plot the data


```{r mice_plot}

# run this once
# get the mice data
# source('class_functions.R')
# download_class_data("mice.Rda")

# load the data
load("mice.Rda")



# plot the data as a boxplot
boxplot(dark_BM_increase, light_BM_increase, 
        names = c("Restricted eating times", "Unrestricted eating times"),
        main = "Comparison of weight gain in mice",
        ylab = "Weight increase (grams")



# plot the data as a strip chart by


# a) creating a list of the data 
data_list <- list(dark_BM_increase, light_BM_increase) 

# b) creating a stripchart   
stripchart(data_list, 
           group.names = c("Dark (restricted)", "Light (unrestricted)"), 
           main = "Comparison of weight gain in mice",
           xlab = "Weight increase (grams)", 
           col = c("red", "blue"))


```




$\\$




##### Step 1: State the null and alternative hypotheses 


$H_0: \mu_{light} - \mu_{dark} = 0$
$H_A: \mu_{light} - \mu_{dark} > 0$


$\alpha = 0.05$





$\\$





##### Step 2: Calculate the observed statistic

```{r mice_obs_stat}


(obs_stat <- mean(light_BM_increase) - mean(dark_BM_increase))


```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_mice}


# combine the data from the dark and light groups together
combined_data <- c(dark_BM_increase, light_BM_increase)


# use a for loop to create shuffled dark and light groups and shuffled statistics 
null_distribution <-  NULL
for (i in 1:10000) {

       # shuffle data
        shuff_data <- sample(combined_data)

        # create fake dark and light groups
        shuff_light   <-  shuff_data[1:9]
        shuff_dark  <-  shuff_data[10:17]

        # save the statistic of interest
        null_distribution[i] <- mean(shuff_light) - mean(shuff_dark)

        
}





```



$\\$





##### Step 4: Calculate a p-value


```{r p_value_mice}


# plot the null distribution again with a red line a the value of the observed statistic
hist(null_distribution, 
     nclass = 200,
     main = "Null distribution", 
     xlab = TeX("$\\bar{x}_{shuff-treat} - \\bar{x}_{shuff-control}$"))


abline(v = obs_stat, col = "red")



# calculate the p-value
(p_value <- sum(null_distribution >= obs_stat)/10000)




```




$\\$






##### Step 5: Make a decision

Since `r p_value` is less than $\alpha = 0.05$ we can reject the null hypothesis and accept the alternative hypothesis. Thus, if we are using the Neyman-Pearson paradigm, we have sufficient evidence to say that eating at night appears to cause weight gain in mice.





$\\$





[Give a mouse a cookie](https://hamstersworld.weebly.com/uploads/2/0/6/1/2061190/8024088.jpg)


[science has gone too far](https://i.imgflip.com/9vr.jpg)



