---
title: "Class 7 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```






$\\$







## Part 1: Parametric tests for comparing two means



**Question:** Does eating late at night make you put on more weight compared to whether you ate the same amount of food during the day? 


To test this idea researchers, [Fonken et al, 2010](https://www.ncbi.nlm.nih.gov/pubmed/20937863), examined this question in mice. 

Mice were randomly divided into 2 groups:

 * **Dark condition**:  8 mice were given 8 hours of darkness at night when they couldn’t eat
 * **Light condition**: 9 were constantly exposed to light for 24 hours so they could always eat


The data from the mice is below. Let's use a t-test to see if there is any difference between the amount of weight the mice gained between these two conditions. 


Our formula for a t-statistic for comparing two means is:  $t = \frac{\bar{x_{light}} - \bar{x_{dark}}}{SE}$ 


Our formula for calculating the standard error is: $SE = \sqrt{\frac{s_1^2}{n_2} + \frac{s_2^2}{n_1}}$


This is Welch's t-test, which gives valid results even when there are unequal variances between the two groups. 





$\\$





##### Step 0: Plot the data


```{r mice_plot}

# run this once
# get the mice data
# source('class_functions.R')
# download_class_data("mice.Rda")

# load the data
load("mice.Rda")



# plot the data as a strip chart by


# a) creating a list of the data 
data_list <- list(dark_BM_increase, light_BM_increase) 

# b) creating a stripchart   
stripchart(data_list, 
           group.names = c("Dark (restricted)", "Light (unrestricted)"), 
           main = "Comparison of weight gain in mice",
           xlab = "Weight increase (grams)", 
           col = c("red", "blue"))


```




##### Step 1: State the null and alternative hypotheses 


$H_0: \mu_{light} - \mu_{dark} = 0$
$H_A: \mu_{light} - \mu_{dark} > 0$


$\alpha = 0.05$





$\\$







##### Step 2: Calculate the observed statistic

```{r pill_obs_stat}


(SE_pooled <- sqrt(     var(light_BM_increase)/length(light_BM_increase) +  var(dark_BM_increase)/length(dark_BM_increase))  )
 
(obs_stat <- (mean(light_BM_increase) - mean(dark_BM_increase))/SE_pooled)


```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_mice}



# sample sizes for the two conditions
n_light <- length(light_BM_increase)
n_dark <- length(dark_BM_increase)


# a conservative estimate of the degrees of freedom
(deg_free <- min(n_light, n_dark) - 1)


# plot the null distribution with the observed statistic on it
x_vals <- seq(-5, 5, length.out = 1000)
y_vals <- dt(x_vals, deg_free)


plot(x_vals, y_vals, type = "l", xlab = "t-statistic values", 
     ylab = "Relative frequency", 
     main = "Null distribution")
points(x_vals, rep(0, length(x_vals)), type = "l")  # add a line at 0



```






$\\$





##### Step 4: Calculate a p-value


```{r p_value_mice}



plot(x_vals, y_vals, type = "l", xlab = "t-statistic values", 
     ylab = "Relative frequency", 
     main = "Null distribution")
points(x_vals, rep(0, length(x_vals)), type = "l")  # add a line at 0
abline(v = obs_stat, col = "red")


# calculate the p-value
(p_value <- pt(obs_stat, df = deg_free, lower.tail = FALSE))




```




$\\$






##### Step 5: Make a decision

Since `r p_value` is less than $\alpha = 0.05$ we can reject the null hypothesis and accept the alternative hypothesis. Thus, if we are using the Neyman-Pearson paradigm, we have sufficient evidence to say that eating at night appears to cause weight gain in mice.







$\\$







